# Cyclistic Bike Share
## About the Company
In 2016, Cyclistic launched a successful bike-share offering company in chicago. Since then, the program has grown to a fleet of 5,824 bicycles that are geotracked and locked into a network of 692 stations across Chicago. The bikes can be unlocked from one station and returned to any other station in the system anytime.

Until now, Cyclistic’s marketing strategy relied on building general awareness and appealing to broad consumer segments.One approach that helped make these things possible was the flexibility of its pricing plans: single-ride passes, full-day passes,and annual memberships. Customers who purchase single-ride or full-day passes are referred to as casual riders. Customers who purchase annual memberships are Cyclistic members.

## Questions for the analysis
* How do annual members and casual riders use Cyclistic bikes differently?
* Why would casual riders buy Cyclistic annual memberships?
* How can Cyclistic use digital media to influence casual riders to become    members?

## In order to answer the key business questions, we will follow:
Data analysis phases: Prepare, Process, Analyze

## Business task
To analyze and identify trends in Cyclistic's historical trip data to help Design marketing strategies build effective campaigns.

# Loading Packages
```{r echo=TRUE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(stringr)
library(dplyr)
```

# Importing Data
```{r echo=TRUE, results='hide',warning=FALSE,message=FALSE}
df1 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202010-divvy-tripdata.csv")
df2 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202011-divvy-tripdata.csv")
df3 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202012-divvy-tripdata.csv")
df4 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202101-divvy-tripdata.csv")
df5 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202102-divvy-tripdata.csv")
df6 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202103-divvy-tripdata.csv")
df7 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202104-divvy-tripdata.csv")
df8 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202105-divvy-tripdata.csv")
df9 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202106-divvy-tripdata.csv")
df10 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202107-divvy-tripdata.csv")
df11 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202108-divvy-tripdata.csv")
df12 <- read_csv("C:/Users/hp/Documents/Cyclistic Bike Share/All Data/202109-divvy-tripdata.csv")
```

# Joining Data
```{r}
bike_rides <- rbind(df1,df2,df3,df4,df5,df6,df7,df8,df9,df10,df11,df12)
bike_rides <- remove_empty(bike_rides,which= c("cols"))
bike_rides <- remove_empty(bike_rides,which= c("rows"))

head(bike_rides)
```


# Cleaning and Processing Data

#### Changing the date formats
```{r}
bike_rides$started_at <- ymd_hms(bike_rides$started_at)
bike_rides$ended_at <- ymd_hms(bike_rides$ended_at)

head(bike_rides)
```

#### Adding more columns Ride length, Day of week, Year, month and Ride duration
```{r}
bike_rides <- bike_rides %>% mutate(Ride_length = seconds_to_period(ended_at-started_at))
bike_rides <- bike_rides %>% mutate(Day_of_week = wday(started_at))
bike_rides <- bike_rides %>% mutate(Year = year(started_at))
bike_rides <- bike_rides %>% mutate(Month = month.abb[month(started_at)])
bike_rides <- bike_rides %>% mutate(Ride_duration = (ended_at-started_at))

head(bike_rides)
```

#### Finding min, max and mean
```{r}
Min_Bike_Rides <- min(bike_rides$Ride_length)
Mean_Bike_Rides <- mean(bike_rides$Ride_length)
Max_Bike_Rides <- max(bike_rides$Ride_length)

Min_Bike_Rides
Mean_Bike_Rides
Max_Bike_Rides
```

## Analysing Data

#### Mean ride length on membership and yearly basis
```{r echo=TRUE,warning=FALSE,message=FALSE}
bike_rides_result <- bike_rides %>% select(member_casual,Ride_length,Year) %>%
group_by(member_casual,Year) %>% summarise(Avg_ride_length = mean(Ride_length))

bike_rides_result
```

#### Number of rides on membership and weekly basis
```{r echo=TRUE,warning=FALSE,message=FALSE}
bike_rides_result1 <- bike_rides %>% select(member_casual,Ride_length,Day_of_week) %>%
group_by(member_casual,Day_of_week) %>% summarise(Number_of_rides = n())

bike_rides_result1
```

#### Number of rides on weekly basis
```{r}
bike_rides_result2 <- bike_rides %>% select(member_casual,Ride_length,Day_of_week) %>%
group_by(Day_of_week) %>% summarise(Number_of_rides = n())

bike_rides_result2
```

#### Summarizing the trip duration
```{r echo=TRUE,warning=FALSE,message=FALSE}
bike_rides_result4 <- bike_rides %>% select(Day_of_week,Ride_duration,member_casual) %>% 
group_by(Day_of_week,member_casual) %>% summarise(Trip_Duration = quantile(minute(seconds_to_period(Ride_duration)),
prob = c(0.15,0.25,0.5,0.75,0.85))) 
```


## Converting Data from Longer to wider format
```{r}
bike_rides_result <- bike_rides_result %>% 
pivot_wider(names_from  = Year, values_from = Avg_ride_length)

bike_rides_result

bike_rides_result1 <- bike_rides_result1 %>% 
pivot_wider(names_from  = Day_of_week, values_from = Number_of_rides)

bike_rides_result1

bike_rides_result2 <- bike_rides_result2 %>% 
pivot_wider(names_from  = Day_of_week, values_from = Number_of_rides)

bike_rides_result2
```

# Visualization

#### Yearly Analysis for Casual vs Annual membership
```{r}
ggplot(data = bike_rides)+ 
geom_bar(mapping= aes(x= Year,fill=member_casual),position="dodge") +
scale_x_continuous(breaks = c(2020,2021))+
facet_wrap(~rideable_type)+
labs(title="Cyclistic Bike Share" , subtitle= "Comparison between Casual and Member Riders Year-On-Year with different Rides", y = "Count of Rides")
```

##### The above analysis might not be appropriate since we have considered last 6 month data out of which we have only 2 months data of year 2022.

#### Monthly Analysis for Casual vs Annual membership
```{r}
ggplot(data = bike_rides)+ 
geom_bar(mapping= aes(x= Month,fill=member_casual)) +
theme(legend.position = "top",axis.text.x = element_text(angle = 45))+ 
labs(title="Cyclistic Bike Share" , subtitle= "Comparison between Casual and Member Riders Month-on-Month basis with different Rides type", y = "Count of Rides")+
facet_wrap(~Year) 
```

```{r}
ggplot(data = bike_rides)+ 
geom_bar(mapping= aes(x= Month,fill=member_casual)) +
theme(legend.position = "top",axis.text.x = element_text(angle = 45))+ 
labs(title="Cyclistic Bike Share" , subtitle= "Comparison between Casual and Member Riders Month-on-Month basis with different Rides type", y = "Count of Rides")+
facet_wrap(~rideable_type)
```


#### Weekly analysis for Casual vs Annual membership
```{r}
ggplot(data = bike_rides)+ 
geom_bar(mapping= aes(x= Day_of_week,fill = factor(Day_of_week))) + theme(legend.position="none")+
facet_wrap(~member_casual) + scale_x_continuous(breaks = c(1:7),
labels = c("Sun","Mon","Tues","Wed","Thurs","Fri","Sat"))+
labs(title="Cyclistic Bike Share" , subtitle= "Analyzing which Week-Day has maximum Rides", y = "Count of Rides")
```

#### Trip duration weekly analysis
```{r}
ggplot(data = bike_rides_result4)+ 
geom_boxplot(mapping= aes(x= factor(Day_of_week), y=Trip_Duration, fill=member_casual),position="dodge") +
theme(legend.position="bottom")+
scale_x_discrete(breaks = c(1:7),labels = c("Sun","Mon","Tues","Wed","Thurs","Fri","Sat")) +
labs(title="Cyclistic Bike Share" , subtitle= "Trip Duration for each Ride on Week-Day basis comparing Casual Vs Annual members", x="Day_of_week" ,y = "Duration (mins)")
```

